---
title: "Female"
author: "Joshua C., Jacob D., Helena GB., Eric W., Brandon H."
group: " Group 5, Section 01"
date: "`r Sys.Date()`"
output: 
  html_document: 
    toc: yes
  pdf_document: 
    latex_engine: xelatex
    toc: yes
  word_document: 
    toc: yes
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, 
                      message = FALSE, 
                      warning = FALSE,
                      comment = "")
```

## Executive Summary

The main goal our group had when analyzing the data was that does the location of the game have any impact of the outcome of match? This is an important question to take into consideration when considering where major games are being held. If a venue is being held closer to one team than the other more of the other teams fans can be there, leading to more of a "home field advantage". When examining this dataset we were hoping to make a correlation so we could prove the statement above. The main questions we were examining to help us prove this were, how does being at home v.s. away affect win probability? Is there a relationship between friendly vs tournament wins/Home_away vs tournament?

Afer performing analysis on our dataset an interesting finding presented itself. That is home teams are 1.33 percent more likely to win. This caught our attention because that it is evidence that helps us more confidently say that venue does play a role on the outcome of the game. We think this matters because sports games are a staple in many world cultures, and outcomes of these games can have strong emotional efffects for the fans. Sports games also have a lot of money surrounding them whether it be betting or a huge income source for that country. 

We recommend that all tournament games be held in a "netural zone". We recommend this for the fact that home teams are more likely to win so to be far to both sides both teams need to have the same advantages.  



Please write your executive summary below. If you prefer, you can type your summary in a text editor, such as Microsoft Word, and paste your final text here.
Include no more than 1/2 page summarizing the context, the primary questions of interest, the data source, the overall approach, and significant findings and/or recommendations.

# Introduction

Rugby is a sport, similar to American football, where an egg-shaped ball is brought to the desired end of the field to score a touchdown. The biggest difference between football and rugby is ironically the fact that in rugby, the ball can only be moved by running and kicking, rather than throwing. This dataset comes from a tidytuesday submission by ScrumQueens, a leader in women’s rugby databases. This data started being collected for fifteens (fifteen players on the field) in 1982 and has been updated for games as recently as May of this year. 

As with many sports, there are so many different ways to analyze how certain factors affect your win percentage, such as being at home, or even just the venue in general. With this dataset, we aim to discover interesting trends relating to how much teams succeed, where they succeed, and if there is any correlation between various factors that go into each game of rugby played.



## Womens Rugby Dataset

```{r}
library(tidyverse)
library(ggplot2)

tuesdata <- tidytuesdayR::tt_load(2022, week = 21)
fifteens <- tuesdata$fifteens
```

### Raw Data
```{r}
fifteens
```

<br>
<br>


# Exploratory Data Analysis


## Fifteens Style Women's Rugby

As data scientists, We have been tasked by the National Rugby Association to look at data from 1982-2022 of all fifteens women rugby matches to see if there are relations between variables. 

The association would like us to look at different venues affect a teams winning to see if any venue has an affect on the outcome. 



### Summary Table

```{r}
#TEMP
ggplot(fifteens, aes(x = venues,)) + geom_bar()
```


### Data Visualization

```{r}

```



## Question 1


**How many distinct teams are in the League?**:


**In the fifteens style of Womens there are 63 countries that have womens teams that compete**



To add additional R code chunks for your work, select `Insert` then `R` from the top of this notebook file or use the keyboard shortcut `Ctrl/cmd` + `Alt/option` + `i`

```{r}
unique(fifteens$team_1)
```


## Question 2



**How many distinct tournaments are there?**:



**There are 39 unique tournaments**:




```{r}
unique(fifteens$tournament)
```



## Question 3



**How many distinct venues are there?**:



**There are 543 unique venues that hosted a rugby game**:


```{r}
unique(fifteens$venue)
```




## Question 4



**How does being at home v.s. away affect win probability**:



**Home teams are 1.33 more likely to win verses being away**:




```{r}
fifteens

table(fifteens$home_away_win)

ggplot(data = fifteens, aes(x = home_away_win)) + geom_bar() + ggtitle("Home Wins vs. Away wins")
```




## Question 5



**Is there a relationship between friendly vs tournament wins?/Home_away vs tournament?**:



**There are a lot less tournament games played which makes for less wins**:




```{r}
ggplot(data = fifteens, aes(x = home_away_win)) + geom_bar() + ggtitle("Home Wins vs. Away Wins")

(wins = FALSE)
ggplot(data = fifteens, aes(x = tournament != "Friendly")) + geom_bar() + ggtitle("Tournament Wins")
```


You can add as many questions as necessary depending on the questions of interests you're trying to address from your dataset (summary and visualizations).

# Summary of Results  

Please write an executive summary of your overall findings and recommendations here. Think of this section as your closing remarks of a presentation, where you summarize your key findings and make recommendations to improve future work or implementation. 

If you prefer, you can type your summary in a text editor, such as Microsoft Word, and paste your final text here.

When it comes to sports, one of the biggest questions is, "Does the venue affect who wins". Our research and findings have analyzed where these teams do their best and why that might be. This data is extremely important to the team especially when it determines where they place in tournaments. The goal of this is to see if the teams do better at home or away games. 

We decided on 5 questions to give us a solution to our question of, "Does the venue affect who wins"?

  1) **How many distinct teams are in the League?**
  2) **How many distinct tournaments are there?**
  3) **How many distinct venues are there?**
  4) **How does being at home v.s. away affect win probability**
  5) **Is there a relationship between friendly vs tournament wins?/Home_away vs tournament?**
  
Some important findings from our data analysis were that teams are 1.33 percent more likely to win at home when compared to away games. Also from the graphs in question 5, you can see that they win more on the graph. Teams, fans, sponsors, and players are all people who contribute to sports teams. That being said, knowing the teams win probability in specific venues will affect how those factors are played. 

We recommend the team to use these stats and data publicly to show those fans, sponsors, and players where they play the best and why that might be. The impact of knowing this information could cause for a number of things:
  * What team players play for?
  * What venue gets sponsored more so a specific team can play more games there?
  * What team is dominant in one venue over another?
  * Does being home vs away affect who wins?

This data recommendation can make or break who wins tournaments or major money winning opportunities for the team. That being said, we believe that every sport should have this analysis done to benefit the teams, fans, sponsors, and players.

Your executive summary must be written in a [professional tone](https://www.universalclass.com/articles/writing/business-writing/appropriate-tone-in-business-communications.htm), with minimal grammatical errors, and should include the following sections:

1. An introduction where you explain the problem statement and goals of your data analysis

    - What problem(s) is this company/team/group trying to solve? Why are they important to their future success?
  
    - What was the goal of your analysis? What questions were you trying to answer and why do they matter?


2. Highlights and key findings from your Exploratory Data Analysis section 
    - What were the interesting findings from your analysis and **why are they important**?

    - This section is meant to **establish the need for your recommendations** in the following section


3. Your recommendations 
  
    - Each recommendation must be supported by your data analysis results 

    - You must clearly explain why you are making each recommendation and which results from your data analysis support this recommendation

    - You must also describe the potential impact of your recommendation:
      
      - Why is this a good recommendation? 
      
# Conclusion
The **conclusion** should be a brief recap of what you did, why you did it, and what you found. Please be sure to **proofread** and **spellcheck** your work before your initial submission.



\newpage

# Appendix

The **appendix** section should contain any additional code, tables, and graphics that are not explicitly referenced in the narrative of the report.










